Refining character relationships using embeddings of textual units : a case study on Les Misérables

### 2022-10-11 article/paper1.tex
write paper

In this article, the data representation for a textual narrative, divided in $n$ textual units $u_1, \ldots, u_n$, is made through two tables. The first one is well known in the field of textual analysis, and consists in the $(n \times v)$ contingency table $\mathbf{N}$, as represented by Table \ref{cont_table}, where $v$ is the vocabulary size. In this table, each row represents an unit, each column a word, and cells $n_{ij}$ counts the number of times the word $i$ appears in the unit. Using this table typically denotes a \emph{Bag-of-Words} approach in our analyses.

@@

The second table, denoted by $\mathbf{O}$, has a size of $(n \times p)$ where $p$ is the number of \emph{character-based objects} found in the narrative and cells $o_{ij}$ counts occurrences of object $j$ in unit $i$. A character-based object can be loosely defined in order to be flexible for various types of narrative or analyses, but can roughly be seen as a countable recurring narrative entity, containing one or more characters. For example, it can be a character occurrence, a character co-occurrence, an oriented character interaction (e.g. a dialog), or even a particular recurring event containing multiple characters (e.g. a meeting). In this article, we mostly consider character occurrences and pair of characters co-occurrences, as shown in Table \ref{occ_table}.

\begin{table}[h]
	\scriptsize
	\begin{tabular}{|c||c|c|c|c|}
		\hline
		& Cosette & Thénardier & Valjean & ... \\
		\hline
		\hline 
		$u_{101}$ & 66 & 78 & 0 & ... \\
		\hline
		$u_{102}$ & 21 & 26 & 0 & ... \\
		\hline
		$u_{103}$ & 12 & 25 & 0 & ... \\
		\hline
		$u_{104}$ & 12 & 0 & 5 & ... \\
		\hline
	\end{tabular}
	\begin{tabular}{|c||c|c|c|c|}
		\hline
		& ... & Cosette-Thénardier & Cosette-Valjean & ... \\
		\hline
		\hline 
		$u_{101}$  	& ... & 66 & 0 &  ... \\
		\hline
		$u_{102}$ 	& ... & 21 & 0 & ... \\
		\hline
		$u_{103}$  	& ... & 12 & 0 & ... \\
		\hline
		$u_{104}$ 	& ... & 0 & 5 & ... \\
		\hline
	\end{tabular}
	\label{occ_table}
	\caption{A snippet of the character-based objects table $\mathbf{O}$ extracted from \emph{les Misérables}. Rows are chapters, columns are character occurrences (top) and character co-occurrences (bottom), and cell $o_{ij}$ counts the number of times object $j$ appear in chapter $i$. Character co-occurrences are computed here as $o_{ij} = \min(o_{ik}, o_{il})$, where $\{k, l\}$ are characters constituting $j$.}
\end{table}

This data organization already gives an orientation to the subsequent analyses and should be kept in mind of the practitioner. Textual unit are now considered as \emph{individuals} (in the statistical terminology), defined by their \emph{variables} contained in the different columns of both tables. Moreover, subsequent analyses are oriented in searching how the object table $\mathbf{O}$ has an influence over the contingency table $\mathbf{N}$, i.e. searching which words are over-represented or under-represented knowing the character-based objects in the unit. While authors use characters in order to build the narrative, we, to a certain extent, work backward: we are searching how character appearances and interactions in the textual unit act on her/his choice of words. If the extraction method permits it, a practitioner should include all character-based objects which she/he desire to study. Here, for example, the choice to include character co-occurrences along with occurrences is motivated by the fact that we are interested in studying character relationships. A character co-occurrence can roughly be seen as an interaction between two characters, and an interaction between two characters (or more, but higher order interactions are outside the scope of this article) should be considered as an object of its own: this interaction is not necessarily the sum of its parts and can give a particular flavor to the unit.

This data organization also highlight the importance of choosing a proper size for the units. These units should be large enough to contain enough words in order the properly capture the textual specificity of each unit, but not too large, as each unit should ideally captures particularities about one of the character-based objects. Unfortunately, it is impossible to define an ideal size for all types of analysis and this size should be balanced regarding the level of analysis, the text size, the selected character-based objects, and previous knowledge of the studied work.

@@

Various methods can be performed on the contingency table $\mathbf{N}$ in order to extract information from it, such as Sentiment Analysis (REF), Textometry (REF), or even Deep Learning methods (REF?). Here, we make to choice to extract a lower dimensional, numeric representation of each units, in other words, an \emph{embedding of textual units}. \\
In section (REF), these vectors of textual units are used as anchor points in order to also embed character-based objects into the same space. Therefore, it is crucial that an interpretation about the directions, or the regions, of this embedding is possible in order to properly extract information about the localization of character-based objects vectors (the relative position of these vectors is insufficient). For that reason, we focus on embeddings of textual units which also contain \emph{vectors of words}: by examining the positions of character-based objects relatively to word vectors, these objects can be characterized.

@@

The relationship between a particular unit $i$ and a word $j$ is expressed through the scalar product between their vectors $s_{ij} := \mathbf{x}^\top_i \mathbf{w}_j$. A positive (resp. negative) similarity denotes a over-representation (resp. under-representation) of the word $j$ in $i$, which permit to find lists of words characterizing the different units. Observe that in this article, this similarity is rather computed between word and character-based object vectors, as defined in section (REF), as they lie in the same space as unit vectors.

Using CA to embedded units however comes with some limitations. First, note that vectors $\mathbf{x}_1, \ldots, \mathbf{x}_n$ obtained from CA reflect textual unit profile (in terms of words) regarding the mean profile. This analysis is thus \emph{contrastive}: it highlights unit variations among the studied text. This could becomes problematic in order to study a dataset containing multiple texts: units belonging to the same text will have a greater chance to be close from each other, and the displayed variance might be largely composed by differences of style between text. Another limitation with this approach is that the words helping the interpretation of units (and character-based objects) are contained in the studied text. Approaches requiring to study the position of units and objects relatively to a predefined list of words (e.g., friends, enemies, family) might therefore be impossible.

@@

Pre-trained word vectors, based on Word2Vec (REF), Glove (REF), or fastText (REF), have recently received great attention from various fields (REFS). They are generally obtained through a training on a very large corpora, such as Wikipedia (REF) or Common Crawl (REF), and the resulting embedding contains a large quantity of word vectors. As shown by multiple studies (REF), these vectors are placed in order to reflect semantic and syntactic similarities between words, and are used in mutliple applications (REF).

There exists various ways to use these pre-trained word vectors in order to find vectors for a \emph{group of words}, such as sentences (REF), paragraphs (REF), or documents (REF). These vectors are often used to apply a classification or clustering algorithm on the newly embedded objects (REF), or to query information (REF). These methods often use the frequencies of words found in the objects, i.e. a table similar to $\mathbf{N}$, but apply various weighting scheme and normalization in order to reduce the effects of frequent words and to align vectors. In the present article, we use a recent methodology proposed in (REF) as it is compatible with multiple unit size and gives state-of-the-art results in multiple tasks.

More precisely, we use a pre-trained word vectors $\mathbf{w}_1, \ldots, \mathbf{w}_m$ trained on Common Crawl and Wikipedia using fastText.\footnote{\url{ https://fasttext.cc/docs/en/crawl-vectors.html}, accessed September 2022.} For French, the number of word-vectors $m$ is 2 millions and the dimension of vectors are $300$. Textual units vectors $\mathbf{x}_1, \ldots, \mathbf{x}_n$ are obtained through the table $\mathbf{N}$ and with the method detailed in Appendix \ref{wv_details}. The main way to interpret units with this method is to use the cosine similarity between its corresponding vector $\mathbf{x}_i$ and every word-vectors $\mathbf{w}_j$, as defined by $s_{ij} := \frac{\mathbf{x}_i^\top \mathbf{w}_j}{\sqrt{\mathbf{x}_i^\top \mathbf{x}_i \mathbf{w}_j^\top \mathbf{w}_j}}$. The cosine similarity also permits to compare units between themselves.

With the pre-trained word vector method, the unit vectors $\mathbf{x}_1, \ldots, \mathbf{x}_n$ (and character-based objects vectors in section REF) lie in an \emph{absolute space} defined by the used word vectors. Comparison between different texts are therefore more pertinent in this space and comparison with words absent from the corpus can be made. However, it is possible that all units from a given text, if the author style is particularly marked, are located in the same region of the space. In this case, the list of most associated word-vectors might be similar for every unit, and the analysis might not give usable results. This effect is fortunately limited by the centration of unit-vectors which occurs in the method describe in Appendix \ref{wv_details}.

\subsection{Character-based object embeddings}

The main goal of this article is not to analyze units, but rather the character-based objects, i.e. the $p$ columns of table $\mathbf{O}$. While we used the table $\mathbf{N}$ to build our embeddings of units, we now use the table $\mathbf{O}$ in order to build the character-based vectors $\mathbf{y}_1, \ldots, \mathbf{y}_p$ which lie in the same space as units.

Two propositions of method are made: the \emph{centroids} method (CENT) is describe in section \ref{centroid} and the \emph{regressions} method (REG) is explained in section \ref{regression}. Both methods work with the proposed embeddings of units.

@@

This method is the most trivial and is based on the following intuition: a character-based object is characterized by all units in which it appears. In other words, we can define the vector $\mathbf{y}_k$ for character-object object $k$ as
\begin{equation}
\mathbf{y}_k = \sum_{i=1}^n \frac{o_{ik}}{o_{\bullet k}} \mathbf{x}_i
\end{equation}
i.e. the center of mass, or \emph{centroid}, of the units containing the character-based object, weighted by the relative weight of its occurrences $\frac{o_{ik}}{o_{\bullet k}}$. This way of building character-based vectors is closely related to the treatment of \emph{supplementary variables} found in CA (REF): these variables do not act in the choice of factorial axis, but can still be represented afterward. However, by contrast, character-based vectors are not dilated after computing centroids, which means that they lie in the same space as units (row). \\
An important remark about the centroid method, is that character-objects vectors positions are \emph{additive}, which means that we have
\begin{equation}
o_{ik} = \sum_{g \in \mathcal{G}} o_{ig}, \forall i \implies \mathbf{y}_k = \sum_{g \in \mathcal{G}} \frac{o_{\bullet g}}{o_{\bullet k}} \mathbf{y}_g,
\end{equation}
where $\mathcal{G}$ is a subset of character-based objects. This property can be interpreted as followed: if the occurrences of a character $k$ can be divided among different situations $g$ (the character alone, the character in interaction with another character, etc.), the character vector $\mathbf{y}_k$ is in fact the centroid of all vectors $\mathbf{y}_g$ of these situations. This is not necessary an undesirable property, but this implies that the specificities of the character might be hidden if he is often registered in an interaction. For example, if the character is frequently with a friend, the character vector will be close to the vector of the couple (if both the character and the couple are counted in the table $\mathbf{O}$). By contrast, if we consider that an interaction between two characters is an \emph{emerging situation}, unrelated to prior behaviors of characters, the regressions method described in the next section seems more appropriate to capture every specificities.

@@

When building a regression model with multiple explanatory variables, it is possible to include every variables and their \emph{interactions}. By doing so, we suppose that the effect of raising both variables is not the same as raising each variable independently. Regression models seem therefore appropriate to capture specificities of having a particular character-based object in a textual unit: the presence of a character $a$ has a effect on the vocabulary of an unit, the presence of another character $b$ has an other effect, and the presence of the pair $\{a, b\}$ yet a different effect. Now, the dependent variable still need to be defined. In fact, we are doing $d$ regressions, with $d$ the number of dimensions of the chosen unit embeddings, and each regressions is constructed to predict the $\alpha$-th coordinate of each units by using variables in the $\mathbf{O}$. In a matrix notation, all models can be written as
\begin{align}
\mathbf{X} = \widetilde{\mathbf{O}} \mathbf{B} + \mathbf{E}, \label{cent_sol}
\end{align}
where $\mathbf{X} = (x_{i\alpha})$ is the $(n \times d)$ matrix containing unit-vectors (on rows), $\widetilde{\mathbf{O}}$ the $(n \times (p+1))$ matrix representing $\mathbf{O}$ with an additional column of $1$ for the intercept, $\mathbf{B} = (\beta_{k\alpha})$ is the $((p+1) \times d)$ matrix containing intercepts and regression coefficients (each regression is contained on a column), and $\mathbf{E}$ the $(n \times \alpha)$ error matrix.

Estimates $\widehat{\mathbf{B}} = (\widehat{\beta}_{k\alpha})$ for the intercept and coefficient regressions are in fact the researched embeddings for character-based objects as well as for the intercept representing the general tone of the studied text. We therefore denote these estimates with $\mathbf{Y} = (y_{k\alpha})$ in the following, with the convention $y_{0\alpha}$ for intercept coefficients.

As the number of character-objects might be very large, it is a good idea to add a $L^2$ regularization term in the objective function. Moreover, the quadratic error rate should also be weighted by the number of tokens in each unit. Including all this, we find the solution for our character-based vectors $\mathbf{y}_0, \mathbf{y}_1, \ldots, \mathbf{y}_p$ contained in rows of $\mathbf{Y}$ with 
\begin{equation}
\mathbf{Y} = (\widetilde{\mathbf{O}}^\top \textbf{Diag}(\mathbf{f}) \widetilde{\mathbf{O}} + \lambda \mathbf{I}_{(p+1)})^{-1} \widetilde{\mathbf{O}}^\top \textbf{Diag}(\mathbf{f}) \mathbf{X}, \label{reg_sol}
\end{equation}
where $\textbf{Diag}(\mathbf{f})$ is the diagonal matrix containing weights of units $\mathbf{f} = \left( \frac{n_{i \bullet}}{n_{\bullet \bullet}} \right)$, $\lambda > 1$ is the regularization coefficient, and $\mathbf{I}_{(p+1)}$ is the identity matrix of size $((p+1) \times (p+1))$.

An interesting effect of the regularization coefficient is that if $\lambda$ is high, equation (\ref{reg_sol}) becomes $\mathbf{Y} \approx \frac{1}{\lambda} \widetilde{\mathbf{O}}^\top \textbf{Diag}(\mathbf{f}) \mathbf{X}$, which is similar to equation (\ref{cent_sol}) (with a contraction and a different weighting scheme). In fact, the regressions method with a regularized term interpolate between the hypothesis where we suppose that every character-object should be considered independently (with $\lambda \to 0$), to the hypothesis of additive mixture between character-objects (with $\lambda \to \infty$), as discussed in section \ref{centroid}. Choosing an appropriate $\lambda$ according to the study (how is another, difficult question) might lead to a situation revealing pertinent information about all character-objects.

@@

For the moment, it is not possible to evaluate the presented methods by the use of some measurements, which would allow to test its validity on various corpora. In order to see if the methods give coherent results, we have to carefully scrutinized and compared them with previous knowledge of the studied work. For this reason, and because of method variations and multiplicity of the results (and lack of place), we chose to present here only one case studies: \emph{Les Misérables} by Victor Hugo. The choice of this work is motivated by the fact that it is a large corpus, immensely studied, and containing various colorful characters and characters relationships. It is therefore a solid choice to clearly illustrate the potential of presented methods.

@@

Characters presences were detected using \emph{Flair}\footnote{\url{https://github.com/flairNLP/flair}} (REF) NER tools. In order to unify character and to further refine the detected characters list, we used hand-made lists of character and aliases from NER results. It resulted in the detection of 54 characters. The character-based objects considered in table $\mathbf{O}$ (Figure \ref{occ_table}) are single character occurrences (54 objects) and character pair co-occurrences (547 objects), resulting in a table of size $365 \times 601$. The character co-occurrences count is computed as $o_{ij} = \min(o_{ik}, o_{il})$, where $\{k, l\}$ are characters constituting the pair $j$.

Note that, in section \ref{diachronic}, we also tested experiments with objects consisting in character and character-pairs as found in each volumes (e.g. Cosette-Valjean in volume one and Cosette-Valjean in volume two are now two different objects) with the addition of volume constants ($V_i=1$ in volume $i$, $V_i=0$ in the other volumes) in order to isolate specific volume vocabulary. This new table $\mathbf{O}_\text{vol}$, containing $1124$ character-based objects, permits to see a diachronic evolution of words describing volumes, characters, and character relationships.

@@

There are two types of methods for unit embeddings, \textbf{CA} (section \ref{ca_method}) and \textbf{WV} (section \ref{wv_method}), as well as two methods for character-based embedding, \textbf{CENT} (section \ref{centroid}) and \textbf{REG} (section \ref{regression}), making a total of 4 possibles ways for obtaining final embeddings. The \textbf{CA} method is fully automated, and results in vectors in a $364$-dimensional space, while the $\textbf{WV}$ methods is based on French \emph{fastText}\footnote{\url{https://fasttext.cc/docs/en/crawl-vectors.html}} (REF) pre-trained embeddings, giving a $300$-dimensional space. The \textbf{CENT} do not have any hyperparameter and we fixed the hyperparameter of the \textbf{REG} methods to $\lambda = 0.01$ for all experiments. \\
Note that, in addition to having two table 4 variations for the methods, 2 tables $\mathbf{O}$ and $\mathbf{O}_\text{vol}$, and a considerable number of characters and character relationships, results can also be presented in various ways (similarities between character-based objects, between words, between objects and words, etc.). Thus, we chose to show here a selection of results for the each method: the 5 most associated words regarding character-based objects (section \ref{words}), the 5 most related character-based object regarding a subset of words (section \ref{objects})), and a diachronic study of top associated words for a subset of character-based objects (section \ref{diachronic}). We invite curious readers to consult these results for all words and objects, which can be found in our GitHub repository.\footnote{\url{https://github.com/gguex/char2char_vectors/results}}

@@

\subsubsection{The most associated words for a subset of characters-based objects}
\label{words}

@@

	\label{word_vs_obj}
	\caption{The 5 most associated words (similarity in parentheses) to a selected set of characters and character pairs, regarding \textbf{CA-CENT}, \textbf{CA-REG}, \textbf{WV-CENT}, and \textbf{WV-REG} methods. Words appearing at least two times within the same method are in bold.}
\end{table*}

@@

The first results in this section is to examine the most associated words with a subset of character-based objects, as measured by the similarities defined in section \ref{unit_embeddings}. Results can be found in Table 3 for all methods.

We can first observe that \textbf{CA} methods seems to summarize objects with a vocabulary closer to the work, while \textbf{WV} methods tend to use more general concepts, with notably more verbs. This behavior can be traced back to the nature of embeddings: in the \textbf{WV} embedding, word-vectors are fixed and do not consider the actual frequencies of words found in the studied corpora. A character could be highly similar to a word appearing only a few times in the corpus if this word is located near the vocabulary associated with this character. By contrast, \textbf{CA} will generally takes into account word frequencies, and balance them with specificities, in order to describe a character.

>>>>

In this article, the data representation for a textual narrative, divided in $n$ textual units $u_1, \ldots, u_n$, is made through two tables. The first one is well known in the field of textual analysis, and consists in the $(n \times v)$ contingency table $\mathbf{N}$, as represented by Table 1, where $v$ is the vocabulary size. In this table, each row represents an unit, each column a word, and cells $n_{ij}$ counts the number of times the word $i$ appears in the unit. Using this table typically denotes a \emph{Bag-of-Words} approach in our analyses.

@@

The second table, denoted by $\mathbf{O}$, has a size of $(n \times p)$ where $p$ is the number of \emph{entities} found in the narrative and cells $o_{ij}$ indicates the presence, or the count for a weighted version, of entity $j$ in unit $i$. An entity, in the context of this article, can be loosely defined in order to be flexible for various types of narrative or analyses, but can roughly be seen as a recurring narrative object, which can contain one or more characters. For example, it can be a character, pair of characters (or even triplet, quadruplet, etc.), an oriented character interaction (e.g. a dialog), or even a particular recurring event containing multiple characters (e.g. a meeting). In this article, we mostly consider character and pair of characters as entities, as shown in Table 2. Note that we consider that a character or a pair of characters are present in the unit if character names (or aliases) are detected above a fixed threshold. A weighted version of this table, where $o_{ij}$ contain the number of occurrences of the entity $j$ in the unit $i$, is also possible and the theory will be written accordingly.

\begin{table}[h]
	\scriptsize
	\begin{tabular}{|c||c|c|c|c|}
		\hline
		& Cosette & Thénardier & Valjean & ... \\
		\hline
		\hline 
		$u_{101}$ & 1 & 1 & 0 & ... \\
		\hline
		$u_{102}$ & 1 & 1 & 0 & ... \\
		\hline
		$u_{103}$ & 1 & 1 & 0 & ... \\
		\hline
		$u_{104}$ & 1 & 0 & 1 & ... \\
		\hline
	\end{tabular}
	\begin{tabular}{|c||c|c|c|c|}
		\hline
		& ... & Cosette-Thénardier & Cosette-Valjean & ... \\
		\hline
		\hline 
		$u_{101}$  	& ... & 1 & 0 &  ... \\
		\hline
		$u_{102}$ 	& ... & 1 & 0 & ... \\
		\hline
		$u_{103}$  	& ... & 1 & 0 & ... \\
		\hline
		$u_{104}$ 	& ... & 0 & 1 & ... \\
		\hline
	\end{tabular}
	\label{occ_table}
	\caption{A snippet of the entities table $\mathbf{O}$ extracted from \emph{les Misérables}. Rows are chapters, columns are characters (top) and character-pairs (bottom), and cell $o_{ij}$ denotes if $j$ appear in chapter $i$.}
\end{table}

This data organization already gives an orientation to the subsequent analyses and should be kept in mind of the practitioner. Textual unit are now considered as \emph{individuals} (in the statistical terminology), defined by their \emph{variables} contained in the different columns of both tables. Moreover, subsequent analyses are oriented in searching how the entity table $\mathbf{O}$ has an influence over the contingency table $\mathbf{N}$, i.e. searching which words are over-represented or under-represented knowing entities in the unit. While authors use characters in order to build the narrative, we, to a certain extent, work backward: we are searching how character appearances and interactions in the textual unit act on her/his choice of words. If the extraction method permits it, a practitioner should include all entities which she/he desire to study. Here, for example, the choice to include character-pairs along with characters is motivated by the fact that we are interested in studying character relationships. A character-pair can roughly be seen as an interaction between two characters, and this interaction should be considered as an object of its own: the presence of this interaction in a unit do not result in having a mixture of words used for each character, and rather give a particular flavor to the unit.

This data organization also highlight the importance of choosing a proper size for the units. These units should be large enough to contain enough words in order the properly capture the textual specificity of each unit, but not too large, as each unit should ideally captures particularities about one of the entity. Unfortunately, it is impossible to define an ideal size for all types of analysis and this size should be balanced regarding the level of analysis, the text size, the selected entities, and previous knowledge of the studied work.

@@

Various methods can be performed on the contingency table $\mathbf{N}$ in order to extract information from it, such as Sentiment Analysis (REF), Textometry (REF), or even Deep Learning methods (REF?). Here, we make to choice to extract a lower dimensional, numeric representation of each units, in other words, an \emph{embedding of textual units}. \\
In section \ref{entity_embeddings}, these vectors of textual units are used as anchor points in order to also embed entities into the same space. Therefore, it is crucial that an interpretation about the directions, or the regions, of this embedding is possible in order to properly extract information about the localization of entity vectors (the relative position of these entity vectors among themselves is insufficient). For that reason, we focus on embeddings of textual units which also contain \emph{vectors of words}: by examining the positions of entities relatively to word vectors, these entities can be characterized.

@@

The association between a particular unit $i$ and a word $j$ is expressed through the scalar product between their vectors $s_{ij} := \mathbf{x}^\top_i \mathbf{w}_j$, which act as a similarity measure. A positive (resp. negative) similarity denotes a over-representation (resp. under-representation) of the word $j$ in $i$, which permit to find lists of words characterizing the different units. Observe that in this article, this similarity is rather computed between word and entity vectors, as defined in section \ref{entity_embeddings}, as they lie in the same space as unit vectors.

Using CA to embedded units however comes with limitations. First, note that vectors $\mathbf{x}_1, \ldots, \mathbf{x}_n$ obtained from CA reflect textual unit profile (in terms of words) regarding the mean profile. This analysis is thus \emph{contrastive}: it highlights unit variations among the studied text. This could becomes problematic in order to study a dataset containing multiple corpora: units belonging to the same text will have a greater chance to be close from each other, and the displayed variance might be largely composed by differences of style between the different works. Another limitation with this approach is that the words helping the interpretation of units (and entities) are contained in the studied text. Approaches requiring to study the position of units and entities relatively to a predefined list of words (e.g., friends, enemies, family) might therefore be impossible if these words do not appear in the text.

@@

Pre-trained word vectors, based on methods such as Word2Vec (REF), Glove (REF), or fastText (REF), have recently received great attention from various fields (REFS). They are generally obtained through a training on a very large corpora, such as Wikipedia (REF) or Common Crawl (REF), and the resulting embedding contains a large quantity of word vectors. As shown by multiple studies (REF), these vectors are placed in order to reflect semantic and syntactic similarities between words, and are used in mutliple applications (REF).

There exists various ways to use these pre-trained word vectors in order to find vectors for a \emph{group of words}, such as sentences (REF), paragraphs (REF), or documents (REF). These vectors are often used to apply a classification or clustering algorithm on the newly embedded objects (REF), or to query information (REF). These methods often use the frequencies of words found in the objects, i.e. a table similar to $\mathbf{N}$, but apply various weighting scheme and normalization in order to reduce the effects of frequent words and to align vectors. In the present article, we use a recent methodology proposed in (REF) as it is compatible with multiple unit sizes and gives state-of-the-art results in various tasks.

Textual units vectors $\mathbf{x}_1, \ldots, \mathbf{x}_n$ are obtained through the table $\mathbf{N}$ and with the method detailed in Appendix \ref{wv_details}. The main way to interpret units with this method is to use the cosine similarity between its corresponding vector $\mathbf{x}_i$ and every word-vectors $\mathbf{w}_j$, as defined by $s_{ij} := \frac{\mathbf{x}_i^\top \mathbf{w}_j}{\sqrt{\mathbf{x}_i^\top \mathbf{x}_i \mathbf{w}_j^\top \mathbf{w}_j}}$. The cosine similarity also permits to compare units between themselves.

With the pre-trained word vector method, the unit vectors $\mathbf{x}_1, \ldots, \mathbf{x}_n$ (and entity vectors in section \ref{entity_embeddings}) lie in an \emph{absolute space} defined by the used word vectors. Comparison between different texts are therefore more pertinent in this space and comparison with words absent from the corpus can be made. However, it is possible that all units from a given text, if the author style is particularly marked, are located in the same region of the space. In this case, the list of most associated word-vectors might be similar for every unit, and the analysis might not give usable results. This effect is fortunately limited by the centration of unit-vectors which occurs in the method describe in Appendix \ref{wv_details}.

\subsection{Entity embeddings}
\label{entity_embeddings}

The main goal of this article is not to analyze units, but rather the entities, i.e. the $p$ columns of table $\mathbf{O}$. While we used the table $\mathbf{N}$ to build our embeddings of units, we now use the table $\mathbf{O}$ in order to build the entity vectors $\mathbf{y}_1, \ldots, \mathbf{y}_p$ which lie in the same space as units.

Two propositions of method are made: the \emph{centroids} method (CENT) is describe in section \ref{centroid} and the \emph{regressions} method (REG) is explained in section \ref{regression}. Both methods work with the aforementioned embeddings of units.

@@

This method is the most trivial and is based on the following intuition: an entity is characterized by all units in which it appears. In other words, we can define the vector $\mathbf{y}_k$ for character-object object $k$ as
\begin{equation}
\mathbf{y}_k = \sum_{i=1}^n \frac{o_{ik}}{o_{\bullet k}} \mathbf{x}_i
\end{equation}
i.e. the center of mass, or \emph{centroid}, of the units containing the entity, weighted by the relative weight $\frac{o_{ik}}{o_{\bullet k}}$. This way of building entity vectors is closely related to the treatment of \emph{supplementary variables} found in CA (REF): these variables do not act in the choice of factorial axis, but can still be represented afterward. However, by contrast, entity vectors are not dilated after computing centroids, which means that they lie in the same space as units (row). \\
An important remark about the centroid method, is that entity vectors positions are \emph{additive}, which means that we have
\begin{equation}
o_{ik} = \sum_{g \in \mathcal{G}} o_{ig}, \forall i \implies \mathbf{y}_k = \sum_{g \in \mathcal{G}} \frac{o_{\bullet g}}{o_{\bullet k}} \mathbf{y}_g,
\end{equation}
where $\mathcal{G}$ is a subset of entities. This property can be interpreted as followed: if a character $k$ can be divided among different situations $g$ (the character alone, the character in interaction with another character, etc.), the character vector $\mathbf{y}_k$ is in fact the centroid of all vectors $\mathbf{y}_g$ of these situations. This is not necessary an undesirable property, but it implies that the specificities of the lone character might be hidden if he is often registered in an interaction. By contrast, if we consider that an interaction between two characters is an \emph{emerging situation}, unrelated to prior behaviors of characters, the regressions method described in the next section seems more appropriate.

@@

When building a regression model with multiple explanatory variables, it is possible to include every variables and their \emph{interactions}. By doing so, we suppose that the effect of raising both variables is not the same as raising each variable independently. Regression models seem therefore appropriate to capture specificities of having a particular entity in a textual unit: the presence of a character $a$ has a effect on the vocabulary of an unit, the presence of another character $b$ has an other effect, and the presence of the pair $\{a, b\}$ yet a different effect. Now, the dependent variable still need to be defined. In fact, we are doing $d$ regressions, with $d$ the number of dimensions of the chosen unit embeddings, and each regressions is constructed to predict the $\alpha$-th coordinate of each units by using variables in the table $\mathbf{O}$. In a matrix notation, all models can be written as
\begin{align}
\mathbf{X} = \widetilde{\mathbf{O}} \mathbf{B} + \mathbf{E}, \label{cent_sol}
\end{align}
where $\mathbf{X} = (x_{i\alpha})$ is the $(n \times d)$ matrix containing unit-vectors (on rows), $\widetilde{\mathbf{O}}$ is the matrix $\mathbf{O}$ with an first additional column of $1$ for the intercept, $\mathbf{B} = (\beta_{k\alpha})$ is the $((p+1) \times d)$ matrix containing intercepts and regression coefficients (each regression is contained on a column), and $\mathbf{E}$ the $(n \times \alpha)$ matrix containing normal errors.

Estimates $\widehat{\mathbf{B}} = (\widehat{\beta}_{k\alpha})$ for the intercept and coefficient regressions are in fact the researched embeddings for entities as well as for the intercept representing the general tone of the studied text. We therefore denote these estimates with $\mathbf{Y} = (y_{k\alpha})$ in the following, with the convention $y_{0\alpha}$ for intercept coefficients.

As the number of entities (i.e. predictors) might be very large, it is a good idea to add a $L^2$ regularization term in the objective function. Moreover, the quadratic error rate should also be weighted by the number of tokens in each unit. Including all this, we find the solution for our character-based vectors $\mathbf{y}_0, \mathbf{y}_1, \ldots, \mathbf{y}_p$ contained in rows of $\mathbf{Y}$ with 
\begin{equation}
\mathbf{Y} = (\widetilde{\mathbf{O}}^\top \textbf{Diag}(\mathbf{f}) \widetilde{\mathbf{O}} + \lambda \mathbf{I}_{(p+1)})^{-1} \widetilde{\mathbf{O}}^\top \textbf{Diag}(\mathbf{f}) \mathbf{X}, \label{reg_sol}
\end{equation}
where $\textbf{Diag}(\mathbf{f})$ is the diagonal matrix containing weights of units $\mathbf{f} = \left( \frac{n_{i \bullet}}{n_{\bullet \bullet}} \right)$, $\lambda > 1$ is the regularization coefficient, and $\mathbf{I}_{(p+1)}$ is the identity matrix of size $((p+1) \times (p+1))$.

An interesting effect of the regularization coefficient is that if $\lambda$ is high, equation (\ref{reg_sol}) becomes $\mathbf{Y} \approx \frac{1}{\lambda} \widetilde{\mathbf{O}}^\top \textbf{Diag}(\mathbf{f}) \mathbf{X}$, which is similar to equation (\ref{cent_sol}) (with a contraction and a different weighting scheme). In fact, the regressions method with a regularized term interpolate between the hypothesis where we suppose that every entity should be considered independently (with $\lambda \to 0$), to the hypothesis of additive mixture between entities (with $\lambda \to \infty$), as discussed in section \ref{centroid}. Choosing an appropriate $\lambda$ according to the study (how is another, difficult question) might lead to a situation revealing desirable information about entities.

@@

For the moment, it is not possible to evaluate the presented methods by the use of some measurements, which would allow to test its validity on various corpora. In order to see if the methods give coherent results, we have to carefully scrutinized and compared them with previous knowledge of the studied work. For this reason, and because of method variations and multiplicity of the results (and lack of place), we chose to present here only one case studies: \emph{Les Misérables} by Victor Hugo. The choice of this work is motivated by the fact that it is a large corpus, well-known, immensely studied, and containing various colorful characters and characters relationships. It is therefore a solid choice to clearly illustrate the potential of presented methods.

@@

Characters were detected using \emph{Flair}\footnote{\url{https://github.com/flairNLP/flair}} (REF) NER tools. In order to unify character and to further refine the detected characters list, we used hand-made lists of character and aliases from NER results. It resulted in the detection of 54 characters. The entities considered in table $\mathbf{O}$ (Figure \ref{occ_table}) are single character (54 entities) and character pairs (547 entities), resulting in a table of size $365 \times 601$. A character or a character-pairs presence is considered present iff characters are detected at least 2 times in the chapter.

Note that, in section \ref{diachronic}, we also tested experiments with entities consisting in character and character-pairs as found in each volumes (e.g. Cosette-Valjean in volume one and Cosette-Valjean in volume two are now two different objects), with the addition of volume constants ($V_i=1$ in volume $i$ and $V_i=0$ in other volumes) in order to isolate specific volume vocabulary. This new table $\mathbf{O}_\text{vol}$, containing $1124$ entities, permits to see a diachronic evolution of words describing volumes, characters, and character relationships.

@@

There are two types of methods for unit embeddings, \textbf{CA} (section \ref{ca_method}) and \textbf{WV} (section \ref{wv_method}), as well as two methods for entity embeddings, \textbf{CENT} (section \ref{centroid}) and \textbf{REG} (section \ref{regression}), making a total of 4 possibles ways for obtaining final embeddings.

The \textbf{CA} method is fully automated, and results in vectors in a $364$-dimensional space, while the $\textbf{WV}$ methods is based on pre-trained word vectors using \emph{fastText} (REF) trained on Common Crawl.\footnote{\url{ https://fasttext.cc/docs/en/crawl-vectors.html}, accessed September 2022.} For French, the number of word-vectors $m$ is 2 millions and the dimension of vectors are $300$.

Note that, in addition to having two table 4 variations for the methods, 2 tables $\mathbf{O}$ and $\mathbf{O}_\text{vol}$, and a considerable number of entities, results can also be presented in various ways (similarities between entities, between entities and words, etc.). Thus, we chose to show here a selection of results for the each method: the 5 most associated words regarding entities (section \ref{words}), the 5 most associated entities regarding a subset of words (section \ref{objects}), and a diachronic study of top associated words for a subset of entities (section \ref{diachronic}). We invite curious readers to consult results for all words and entities, which can be found in our GitHub repository.\footnote{\url{https://github.com/gguex/char2char_vectors/results}}

@@

\subsubsection{The most associated words for a subset of entities}
\label{words}

@@

	\label{word_vs_obj}
	\caption{The 5 most associated words (similarity in parentheses) to a selected set of entities, regarding \textbf{CA-CENT}, \textbf{CA-REG}, \textbf{WV-CENT}, and \textbf{WV-REG} methods. Words appearing at least two times within the same method are in bold.}
\end{table*}

@@

The first results in this section is to examine the most associated words with a subset of entities, as measured by the similarities defined in section \ref{unit_embeddings}. Results can be found in Table 3 for all methods.
